---
title: "LPfars: How it works"
author: "Lefkios Paikousis"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{LPfars: How it works}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This is a vignette that demontrates the use of the 3 main functions of the `LPfars` package

- `fars_read`: Read a single dataset stored on your disk
- `fars_summarize_years`: Summarize the number of observations by year
- `fars_map_state`: Visualize the accidents in the US map

# Read FARS Data

Through `fars_read`, we can read the FARS data that was saved by csv file with ".bz2" extension. 

_NOTE:_ The LPfars package comes with example raw datasets. The location of the example raw datasets on your disk can be accessed by typing

`system.file("extdata", package = "LPfars")`

To access the name of a single file, type    

`system.file("extdata", "the_file_name", package = "LPfars")`

```{r}
library(LPfars)

file <- system.file("extdata", "accident_2013.csv.bz2", package = "LPfars")

data <- fars_read(file)

print(data)

```


# Summary of FARS Data

The `fars_summarize_years` returns a summary table with the number of monthly accident cases (row) and the defined year(s) (column). Here the input parameter canbe a numeric vector or a list of the year values.

```{r}

# Set the working directory to the location of the files
# This is for demontrating purposes. Once you have your datasets in your current working directory, you will not have to do this

files_location = system.file("extdata", package = "LPfars")

setwd(files_location)

fars_summarize_years(2013:2015)

```
_NOTE:_ The function `fars_summarise_years` calls the function `fars_read_years`to read multiple FARS data files on your disk


# Plot the density of accident locations

The `fars_map_state` returns a map with the plot of the state with accident locations. The State index was defined from 1 to 56.

```{r}

files_location = system.file("extdata", package = "LPfars")
setwd(files_location)

# States are recored as integers. Lets have a look at them
unique(data$STATE) 


# Lets pick state # 1 at year 2013

 fars_map_state(1, 2013)


```
